Concrete Semantics for Pushdown Analysis:\\The Essence of Summarization

### 2013-10-08 content.tex
Changed page size to the actual page size used. Updated related work.

\section{What to expect}

CFA2 uses a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
%
Summarization algorithms need not be restricted to languages with well-bracketed calls and returns.
%
We can adopt the technique for higher precision in the common case but still handle difficult cases such as first-class control.
%
This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}.
%
This impressive work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
%
Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, AAM would provide.
%%

@@

There are common underpinnings of PDCFA and CFA2 that can be embodied as concrete machinery in the programming language semantics: 
\begin{enumerate}
\item{breaking the recursive structure of continuations by indirecting them through a table with appropriately precise keys, and}
\item{memoization.}
\end{enumerate}

Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.
%
In \autoref{sec:pdcfa} we derive a cousin of PDCFA.
%
The following section shows how to extend PDCFA with abstract garbage collection. % \autoref{sec:gc}: we show how abstract garbage collection ``falls out'' of this approach.
%
Then in \autoref{sec:cfa2} we show the fundamental difference between PDCFA and CFA2 without first-class control. % : we make additions to the PDCFA semantics to get a direct-style CFA2 without first-class control}
%
\ifwcm{We show in \autoref{sec:wcm} that semantics with stack-inspection (beyond GC) also fit within this interpretation.}
%
Finally, we show in \autoref{sec:sr} that even though first-class control operators don't fit in the pushdown model, we can gracefully loosen the abstraction to analyze delimited, composable first-class control.

@@

%% FIXME: remove stores from memo-table since they will always be unchanged
%% TODO: Show that store widening and store/memo/KTab widening have a bisimulation.
{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{align*}
  \mastate \in \sa{State} &= (\Expr \times \Env) \times \Store \times \Kont \\
  \System &= \wp(\sa{State}) \times \wp(\sa{State}) \times \KTab \times \Memo \\
  {\mathcal F} &: \sa{\State} \to \System \to \System \\
  {\mathcal F}(\mstate_0)(S, F, \mktab, \mmemo) &= (S \cup S', F' \cup (\set{\mstate_0} \setminus S), \mktab', \mmemo') \\
  \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
                \wn(\tpl{(\mexpr,\menv),\mstore,\mkont},\mktab,\mmemo)
                   &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
\end{align*}
\begin{minipage}{.55\linewidth}
  \begin{align*}
    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \_, \mktab', \_) \in I} \\
    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I}
  \end{align*}
\end{minipage}
\begin{minipage}{.40\linewidth}
  \begin{align*}
    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_, \_) \in I} \\
    F' &= S' \setminus S
  \end{align*}
\end{minipage}
}
%%

%%
A $\System$ embodies the states seen and at which store, $S$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
% Note: Only if widened store.{
%
%The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
%
%All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
%
%The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
%
Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
%
PDCFA claims there is no loop, and depending on the implementation, that the result is true or false (paper), or just true (implemented)
\footnote{This is because the paper steps every seen state with the current store every iteration, but the implementation only steps states that need stepping.}.
%
Let's take a look at the evaluation after binding \texttt{id}:

\begin{center}
  \begin{minipage}{0.55\linewidth}
    \begin{center}
      $\begin{array}{l}\includegraphics[scale=0.65]{example1}\end{array}$ \\
      $\Downarrow$ \\
      $\begin{array}{l}\includegraphics[scale=0.65]{example2}\end{array}$ \\
      $\Downarrow$ \\
      $\begin{array}{l}\includegraphics[scale=0.65]{example3}\end{array}$
    \end{center}
  \end{minipage}
  \begin{minipage}{0.43\linewidth}
    \begin{center} {\small
        \begin{tabular}{|l|c|c|}
          \hline
          \# & $\mstore$ & $\mkont$ \\
          \hline
          0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\bot)}]$ & $\mkont_0$ \\
          1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
          2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ \\
          3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
          4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
          \hline
        \end{tabular}} \\[10pt]
      {\small
        \begin{tabular}{|l|c|}
          \hline
          \# & $\mctx$ \\
          \hline
          0  & $(({\tt x}, \iota), \mstore_1)$ \\
          1  & $(({\tt x}, \iota), \mstore_3)$ \\
          \hline
        \end{tabular}}
      \\[10pt]
      {\small
        \begin{tabular}{|l|c|c|c|}
          \hline
          \# & $\mktab$ & $\mmemo$ \\
          \hline
          0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
          1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
          \hline
        \end{tabular}}    
    \end{center}
  \end{minipage}
\end{center}
Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
% %% FIXME stores in memo-table
% \begin{enumerate}
% \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
%       (call this $\mkont_2$).}
% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
% \end{enumerate}

We maintained enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mmemo$ to continue past function calls, so there is no confusion about a back edge from the second call to \texttt{id}.
%

@@

The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined as the greatest fixed point of the following rules:
\begin{mathpar}
  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt}} \quad
  \inferrule{\mkont \in \unroll{\mktab}{\mkont'}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont'}} \quad
  \inferrule{\mkont^\circ \in \mktab(\mctx) \\
             \mkont \in \unroll{\mktab}{\mkont^\circ}}
            {\mkont \in \unroll{\mktab}{\krt{\mctx}}}
\end{mathpar}

The astute reader may find this definition dubious, since there is a possibility for unguarded corecursion here.
%
Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
%
The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.
%
Our interpretation of $\stepto$ is in a fixed-point computation, so circularities are not a problem for termination, but we augment the definition here to guard against circularities in our proofs:

\begin{mathpar}
  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
  \inferrule{\mctx \notin G  \\
             \mkont^\circ \in \mktab(\mctx) \\
             \mkont \in \unroll{\mktab}{\mkont^\circ, \set{\mctx}\cup G}}
            {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
\end{mathpar}

%%
The different tables we have encode information about execution history that we prove is invariant.
\begin{align*}
  \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
          \dom(\mmemo) \subseteq \dom(\mktab) \\
  &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab),
                  \mkont^\circ \in \mktab(\mpoint',\mstore'),
                  \mkont' \in \unroll{\mktab}{\mkont^\circ, \emptyset}. \\   
  &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
  \text{where } \phi_{\reachable} &=
  \exists \mkont^* \in \unroll{\mktab}{\mkont,\emptyset}.
    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
  \phi_{\memo} &=
 \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
  &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mval,\mstore'', \mkont'}.
     \hastail(\mtrace,\mkont') \\
\end{align*}

The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
%
The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
%
The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).

\begin{lemma}[Table invariants]\label{lem:tab-inv}
  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
\end{lemma}

@@

Abstract garbage collection~\citep{dvanhorn:Might:2006:GammaCFA} is a powerful technique for regaining precision in an abstract interpretation.
%
The followup to PDCFA~\citep{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
%
In our view with tabular semantics, the GC story mimics the simplicity of GC in the regular setting, without the detour into automata theory.
%
We follow the GC style of \citet{dvanhorn:Might:2006:GammaCFA} with a stop-and-copy collector.
%
Address reachability is in terms of a base ``touches'' function, $\touches$, that is different for our continuations.
%
Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
\begin{align*}
  \touchesk{\mkont}(\mktab) &= \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^* \maddr}
\\[2pt]
  \touchesf(\kar{\mexpr,\menv}) &= \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)}) \\
  \touchesf(\kfn{\mval}) &= \touches(\mval)
\end{align*}
\begin{mathpar}
  \inferrule{ }{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont}
  \quad
  \inferrule{\maddr \in \touchesf(\mkframe)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
  \quad
  \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
\end{mathpar}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Begin scary Scribble output
\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touch{-}kont}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mktab$}\RktPn{)}

@@

\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{let}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktPn{[}\RktSym{$\mkont$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{]}\RktPn{)}

\mbox{\hphantom{\Scribtexttt{xxxx}}}\RktPn{(}\RktSym{match}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}

@@

\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{cons}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkframe$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\cup$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touches{-}frame}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkframe$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{loop}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\mkont$}\RktPn{)}\RktPn{)}\RktPn{]}

@@

%%
A new temporal logic that itself understands well-bracketing of call and returns for stating and validating propositions, NWTL ~\citep{ianjohnson:alur:nwtl:2007}, applies to visibly pushdown systems, which can precisely model programming languages that only have well-bracketed call and returns.
%
The propositional form of the logic has decidable satisfiability checking, and control-flow queries are shallow propositions to pose in the logic.
%
However, the satisfiability problem is \textsc{2Exptime}-hard and the algorithm is not given in the on-the-fly form that a higher-order language would need to effectively answer control-flow queries without already knowing the answers.
%
The logic itself is an exciting new frontier for expressing program correctness properties, and a higher-order version would be a welcome addition to the functional programmer's tool belt.
%%

>>>>

\section{The essence of summarization}

CFA2 uses a technique called ``summarization'' from Sharir and Pnueli \citep[Chapter 7]{local:muchnick:jones:flow-analysis:1981}, which is synonymous with the $\epsilon$-closure graph that PDCFA constructs.
%
Summarization algorithms need not be restricted to languages with well-bracketed calls and returns.
%
We can adopt the technique for higher precision in the common case but still handle difficult cases such as first-class control.
%
This was shown for the \rackett{call-with-current-continuation} (a.k.a. \rackett{call/cc}) operator in ~\citet{ianjohnson:Vardoulakis2011Pushdown}.
%
This impressive work illuminated the fact that we can harness the enhanced technology of pushdown analyses in non-pushdown models of computation.
%
Doing this sacrifices call/return matching in the general case, but in practice the precision is much better than the alternative regular model that, say, 0-CFA would provide.
%%

@@

There are common underpinnings of PDCFA and CFA2 that can be embodied as concrete ``summarization'' machinery in the programming language semantics: 
\begin{enumerate}
\item{bound continuations with with delimiters that specify the exact context in which any continuation replacement would lead to the same result, and}
\item{memoize at these delimiters}
\end{enumerate}

%%
The first of these two might seem unfamiliar, but it is just identifying what is needed for sound memoization, and we delimit the continuations in order to mark where to memoize results.
%
In the case of a CESK machine, where the store is relevant to the evaluation of functions, a context is just the CES portion since stacks don't matter.
%
For stack inspecting analyses, a context is CES plus the inspected portion of K.
%
As soon as we have a value that such a context evaluates to, all subsequent visits to that context can jump straight to the memoized result(s).
%%

%%
Once the machine semantics is in this form, simple point-wise abstraction leads to the summarization algorithms that we see in the literature.
%
In \autoref{sec:pdcfa} we derive a cousin of PDCFA.
%
The following section shows how to extend PDCFA with abstract garbage collection. % \autoref{sec:gc}: we show how abstract garbage collection ``falls out'' of this approach.
%
Then in \autoref{sec:cfa2} we show the fundamental difference between PDCFA and CFA2 without first-class control. % : we make additions to the PDCFA semantics to get a direct-style CFA2 without first-class control}
%
\ifwcm{We show in \autoref{sec:wcm} that semantics with stack-inspection (beyond GC) also fit within this interpretation.}
%
Finally, we show in \autoref{sec:sr} that even though first-class control operators don't fit in the pushdown model, we can gracefully loosen the abstraction to analyze delimited, composable first-class control.

@@

%% FIXME: remove stores from memo-table since they will always be unchanged
%% TODO: Show that store widening and store/memo/KTab widening have a bisimulation.
{
\setlength{\abovedisplayskip}{0pt}
\setlength{\belowdisplayskip}{4pt}
\setlength{\abovedisplayshortskip}{0pt}
\setlength{\belowdisplayshortskip}{8pt}
\begin{align*}
  \mastate \in \sa{State} &= (\Expr \times \Env) \times \Store \times \Kont \\
  \System &= \wp(\sa{State}) \times \wp(\sa{State}) \times \KTab \times \Memo \\
  {\mathcal F} &: \sa{\State} \to \System \to \System \\
  {\mathcal F}(\mastate_0)(S, F, \mktab, \mmemo) &= (S \cup S', F', \mktab', \mmemo') \\
  \text{where } I &= \setbuild{\mstate'}{\mastate \in F, \wn(\mastate,\mktab,\mmemo) \stepto \mstate'} \\
                \wn(\tpl{(\mexpr,\menv),\mstore,\mkont},\mktab,\mmemo)
                   &= \tpl{(\mexpr,\menv),\mstore,\mkont,\mktab,\mmemo}
\end{align*}
\begin{minipage}{.55\linewidth}
  \begin{align*}
    \mktab' &=  \bigsqcup\setbuild{\mktab'}{\wn(\_, \_, \mktab', \_) \in I} \\
    \mmemo' &= \bigsqcup\setbuild{\mmemo'}{\wn(\_, \_, \_, \mmemo') \in I}
  \end{align*}
\end{minipage}
\begin{minipage}{.40\linewidth}
  \begin{align*}
    S' &= \setbuild{\mastate'}{\wn(\mastate', \_, \_, \_) \in I} \cup \set{\mastate_0}\\
    F' &= S' \setminus S
  \end{align*}
\end{minipage}
}
%%

%%
A $\System$ embodies the states seen and at which store, $S$, the frontier set of states yet to analyze, $F$, the continuation table $\mktab$ and the memo table $\mmemo$.
% Note: Only if widened store.{
%
%The memo table need only map to values, since the global store will always subsume the store that would otherwise be mapped.
%
%All the states in the frontier are stepped with the current store, after which the next store is the least upper bound of all the resulting stores.
%
%The next frontier contains only states resulting from stepping the previous frontier that we haven't seen at this next store.}
%
The least fixed-point of ${\mathcal F}(\tpl{(\mexpr,\bot),\bot,\kmt})$ is the analysis of closed expression $\mexpr$.
%
Systematic techniques for a performant implementation can be found in \citet{ianjohnson:oaam:icfp2013}.

@@

%%
We should expect that a pushdown analysis would predict this evaluates to true, and there are no loops in the program.
%
0CFA~\citep{dvanhorn:Shivers:1991:CFA} claims there is a loop from the second call of \texttt{id} to the first, and thusly predicts this program evaluates to true or false.
%
PDCFA claims there is no loop, and depending on the implementation, that the result is true or false (paper), or just true (implemented)
\footnote{This is because the paper steps every seen state with the current store every iteration, but the implementation only steps states that need stepping.}.
%
The example evaluation is in \autoref{fig:ex-eval}.
%
We maintain enough context to distinguish the return points of \texttt{id} to not rebind \texttt{y} to 1.
%
When determining control flow through the expression, we consult $\mmemo$ to continue past function calls, so there is no confusion about a back edge from the second call to \texttt{id}.
%

\begin{figure}
  \centering
    \begin{minipage}{0.55\linewidth}
      \begin{center}
        $\begin{array}{l}\includegraphics[scale=0.65]{example1}\end{array}$ \\
        $\Downarrow$ \\
        $\begin{array}{l}\includegraphics[scale=0.65]{example2}\end{array}$ \\
        $\Downarrow$ \\
        $\begin{array}{l}\includegraphics[scale=0.65]{example3}\end{array}$
      \end{center}
    \end{minipage}
    \begin{minipage}{0.43\linewidth}
      \begin{center} {\small
        \begin{tabular}{|l|c|c|}
          \hline
          \# & $\mstore$ & $\mkont$ \\
          \hline
          0 & $[{\tt id} \mapsto \set{(\slam{\mvar}{\mvar},\bot)}]$ & $\mkont_0$ \\
          1 & $\mstore_0[{\tt \mvar} \mapsto \set{0}]$ & $\krt{\mctx_0}$ \\
          2 & $\mstore_1[{\tt y} \mapsto \set{0}]$ & $\mathbf{lt}({\tt z}, {\tt (}\le\ {\tt y}\ {\tt z)})$ \\
          3 & $\mstore_2[{\tt x} \mapsto \set{0,1}]$ & $\krt{\mctx_1}$ \\
          4 & $\mstore_3[{\tt z} \mapsto \set{0,1}]$ & $\kmt$ \\
          \hline
        \end{tabular}} \\[10pt]
      {\small
        \begin{tabular}{|l|c|}
          \hline
          \# & $\mctx$ \\
          \hline
          0  & $(({\tt x}, \iota), \mstore_1)$ \\
          1  & $(({\tt x}, \iota), \mstore_3)$ \\
          \hline
        \end{tabular}}
      \\[10pt]
      {\small
        \begin{tabular}{|l|c|c|c|}
          \hline
          \# & $\mktab$ & $\mmemo$ \\
          \hline
          0  & $[\mctx_0 \mapsto \set{\mkont_0}]$ & $[\mctx_0 \mapsto \set{0}]$ \\
          1  & $\mktab_0[\mctx_1 \mapsto \set{\mkont_2}]$ & $\mmemo_0[\mctx_1 \mapsto \set{0,1}]$ \\
          \hline
        \end{tabular}}    
    \end{center}
  \end{minipage}
Where $\mkont_0 = \mathbf{lt}({\tt y}, {\tt (let\ ([z\ (id\ 1)])\ (}\le\ {\tt y}\ {\tt z))}, \iota)$
\caption{Example analysis evaluation}
\label{fig:ex-eval}
\end{figure}
% %% FIXME stores in memo-table
% \begin{enumerate}
% \item{\texttt{(id 0)} steps to \texttt{x} at $\krt{((\texttt{x}, [\texttt{x}\mapsto\texttt{x}]), \mstore_0)}$ with $\mstore_1 =
%     \mstore_0[\texttt{x} \mapsto \set{0}]$ and
%     (let $\mctx_1 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_1)$,
%          $\mkont_1 = \klet{\texttt{y}, \texttt{(let ([z (id 1)]) ($\le$ y z))}, [\texttt{id}\mapsto{\texttt{id}}]}$)
%     $\mktab_1 = [\mctx_0 \mapsto \set{\mkont_0}]$}
% \item{\texttt{0} at $\krt{\mctx_1}$ steps to $0$ at $\mkont_1$ and $\mmemo_1 = [\mctx_1 \mapsto \set{0}]$}
% \item{\texttt{0} at $\mkont_1$ steps to \texttt{(let ([z (id 1)]) ($\le$ y z))} and $\mstore_2 = \mstore_1[\texttt{y} \mapsto \set{0}]$}
% \item{\texttt{(let ([z (id 1)]) ($\le$ y z))} steps to \texttt{(id 1)} at \\
%       $\klet{\texttt{z}, \texttt{($\le$ y z)}, [\texttt{id} \mapsto \texttt{id}, \texttt{y} \mapsto \texttt{y}]}$
%       (call this $\mkont_2$).}
% \item{\texttt{(id 1)} steps to $x$ with $\mstore_3 = \mstore_2[\texttt{x} \mapsto \set{0,1}]$ and
%       (let $\mctx_3 = ((\texttt{x},[\texttt{x}\mapsto\texttt{x}]),\mstore_3)$) $\mktab_3 =\mktab_1[\mctx_3 \mapsto \set{\mkont_2}]$.}
% \item{\texttt{0} or \texttt{1} at $\krt{\mctx_3}$ steps to \texttt{0} or \texttt{1} at $\mkont_2$ and $\mmemo_3 = \mmemo_1[\mctx_3 \mapsto \set{0,1}]$.}
% \item{\texttt{z} gets bound to $\set{0,1}$, and \texttt{($\le$ y z)} evaluates to true.}
% \end{enumerate}

@@

The $\mathit{unroll}$ function interprets what are all the valid continuations that $\mktab$ encodes for a given continuation that contains $\krt{\mctx}$, defined as the greatest fixed point of the following rules:
\begin{mathpar}
  \inferrule{ }{\kmt \in \unroll{\mktab}{\kmt, G}} \quad
  \inferrule{\mkont \in \unroll{\mktab}{\mkont', \emptyset}}{\mkframe:\mkont \in \unroll{\mktab}{\mkframe:\mkont', G}} \\
  \inferrule{\mctx \notin G  \\
             \mkont^\circ \in \mktab(\mctx) \\
             \mkont \in \unroll{\mktab}{\mkont^\circ, \set{\mctx}\cup G}}
            {\mkont \in \unroll{\mktab}{\krt{\mctx}, G}}  
\end{mathpar}

We add $G$ to protect against unguarded corecursion in order for the definition to be well-defined.
%
Interpreting a function with ill-founded recursion can lead to a table such that $\krt{\mctx} \in \mktab(\mctx)$.
%
The abstractions that we impose on our data might also cause this situation, even though the concrete execution might always terminate.

%%
The different tables we have encode information about execution history that we prove is invariant (\autoref{fig:inv}).
%
The first is that the memo table only contains information about previously seen contexts; we need this to infer that there was at least one call leading to the memoized context such that we can use stack irrelevance to justify skipping to the memoized result.
%
The second, $\phi_{\reachable}$, states that all the calling contexts in the continuation table reach some unrolling of the current state.
%
The final invariant, $\phi_{\memo}$, states that the memo table is correct: all calling contexts reach the memoized result, with the calling context's continuation in the tail (to justify using stack-irrelevance).

\begin{figure}
  \centering
  \begin{align*}
    \inv(\tpl{\mpoint, \mstore, \mkont, \mktab, \mmemo}) &=
    \dom(\mmemo) \subseteq \dom(\mktab) \\
    &\wedge \forall (\mpoint',\mstore') \in \dom(\mktab), \mkont^\circ
    \in \mktab(\mpoint',\mstore'),
    \mkont' \in \unroll{\mktab}{\mkont^\circ, \emptyset}. \\
    &\qquad \phi_{\reachable} \wedge \phi_{\memo} \\
    \text{where } \phi_{\reachable} &= \exists \mkont^* \in
    \unroll{\mktab}{\mkont,\emptyset}.
    \tpl{\mpoint', \mstore', \mkont'} \stepto^* \tpl{\mpoint, \mstore, \mkont^*} \\
    \phi_{\memo} &=
    \forall (\mval, \mstore'') \in \mmemo(\mpoint',\mstore'). \\
    &\qquad \exists \mtrace\equiv\tpl{\mpoint', \mstore', \mkont'}
    \stepto^* \tpl{\mval,\mstore'', \mkont'}.
    \hastail(\mtrace,\mkont') \\
  \end{align*}
  \caption{Table invariants}
\label{fig:inv}
\end{figure}
\begin{lemma}[Table invariants]\label{lem:tab-inv}
  If $\inv(\mstate)$ and $\mstate \stepto \mstate'$ then $\inv(\mstate')$.
\end{lemma}

@@

Abstract garbage collection~\citep{dvanhorn:Might:2006:GammaCFA} is a powerful technique for regaining precision in an abstract interpretation.
%
The followup to PDCFA~\citep{dvanhorn:Earl2012Introspective} that allows inspecting the stack in order to perform sound GC introduces an entirely new class of automata for this one purpose.
%
In our view with tabular semantics, the GC story mimics the simplicity of GC in the regular setting, without the detour into automata theory.
%
We follow the GC style of \citet{dvanhorn:Might:2006:GammaCFA} with a stop-and-copy collector.
%
Address reachability is in terms of a base ``touches'' function, $\touches$, that is different for our continuations.
%
Entries in $\mktab$ might be circular, so we compute the continuation frames in a fixed-point before computing touched addresses:
\begin{align*}
  \touchesk{\mkont}(\mktab) &= \setbuild{\maddr}{\mkont \rightsquigarrow_\mktab^* \maddr}
\\[2pt]
  \touchesf(\kar{\mexpr,\menv}) &= \mathit{rng}(\menv|_{\mathit{fv}(\mexpr)}) \\
  \touchesf(\kfn{\mval}) &= \touches(\mval)
\end{align*}
\begin{mathpar}
  \inferrule{ }{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont}
  \quad
  \inferrule{\maddr \in \touchesf(\mkframe)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \mkont'}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \mkont'}
  \quad
  \inferrule{\mkont \rightsquigarrow_\mktab \maddr}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab \maddr}
%  \inferrule{\mkont \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}{\kcons{\mkframe}{\mkont} \rightsquigarrow_\mktab (\mkont' \text{ or } \maddr)}
  \quad
  \inferrule{\mkont \in \mktab(\mctx)}{\krt{\mctx} \rightsquigarrow_\mktab \mkont}
\end{mathpar}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Begin scary Scribble output
\begin{SCodeFlow}\begin{RktBlk}\begin{SingleColumn}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touch{-}kont}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\Xi$}\RktPn{)}

@@

\mbox{\hphantom{\Scribtexttt{xx}}}\RktPn{(}\RktSym{define}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}

\mbox{\hphantom{\Scribtexttt{xxxx}}}\RktPn{(}\RktSym{match}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}

@@

\mbox{\hphantom{\Scribtexttt{xxxxxx}}}\RktPn{[}\RktPn{(}\RktSym{cons}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\phi$}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{$\cup$}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{touches{-}frame}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\phi$}\RktPn{)}\mbox{\hphantom{\Scribtexttt{x}}}\RktPn{(}\RktSym{fix{-}touch}\mbox{\hphantom{\Scribtexttt{x}}}\RktSym{$\kappa$}\RktPn{)}\RktPn{)}\RktPn{]}

@@
